---
title: "Assignment04"
author: "Yuwei Wu"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(microbenchmark)
library(Matrix)
library(parallel)
library(tinytex)
library(timeSeries)
```

# HPC
## Problem 1: Make sure your code is nice
Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google
```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  # YOUR CODE HERE
  ans<-rowSums(mat)
  ans
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  # YOUR CODE HERE
  ans <- t(apply(mat, 1, cumsum))
  ans
}

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
first <- microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), check = "equivalent"
)
print(first, unit = "relative")

# Test for the second
sec <- microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), check = "equivalent"
)
print(sec, unit = "relative")
```

The last argument, check = “equivalent”, is included to make sure that the functions return the same result.

## Problem 2: Make things run faster with parallel computing
The following function allows simulating PI
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:
```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():
```{r}
# YOUR CODE HERE
cl <- makePSOCKcluster(4L) 
clusterSetRNGStream(cl, 123)
clusterExport(cl, c("sim_pi"), envir = environment())

system.time({
  ans <- unlist(parLapply(cl,1:4000, sim_pi, n = 10000)) 
  print(mean(ans))
})

stopCluster(cl)
```


# SQL
Setup a temporary database by running the following chunk
```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```


## Question 1
How many many movies is there avaliable in each rating catagory.
```{sql connection=con}
SELECT rating,
  COUNT(*) AS count
FROM film
GROUP BY rating
```

There are 180 movies with the rating "G"; 210 movies with the rating "NC-17"; 194 movies with the rating "PG"; 233 movies with the rating "PG-13"; 195 movies with the rating "R".

## Question 2
What is the average replacement cost and rental rate for each rating category.
```{sql, connection = con}
SELECT rating,
  AVG(replacement_cost) AS avg_replcement,
  AVG(rental_rate) AS avg_rental
FROM film
GROUP BY rating
```

For the rating "G", the average replacement cost is 20.12 and the average rental rate is 2.91. 
For the rating "NC-17", the average replacement cost is 20.13 and the average rental rate is 2.97. 
For the rating "PG", the average replacement cost is 18.96 and the average rental rate is 3.05.
For the rating "PG-13", the average replacement cost is 20.40 and the average rental rate is 3.03.
For the rating "R", the average replacement cost is 20.23 and the average rental rate is 2.94.

## Question 3
Use table film_category together with film to find the how many films there are witth each category ID
```{sql, connection = con}
SELECT category_id,
  COUNT (*) AS Counts
FROM film AS f
  INNER JOIN film_category AS c
ON f.film_id = c.film_id
GROUP BY category_id
```

## Question 4
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql, connection = con}
SELECT film_category.category_id,category.name,
  COUNT(*) AS count
FROM film_category
  INNER JOIN film ON film_category.film_id=film.film_id
  INNER JOIN category ON film_category.category_id=category.category_id
GROUP BY category.category_id
ORDER BY count DESC
```

The most popular category is sports. 